---
title: "moredynamic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(here)

source(here::here("Week6/dexppop.R"))
source(here::here("Week6/exppop.R"))
source(here::here("Week6/dexppop_play.R"))
source(here::here("Week6/discrete_logistic_pop.R"))
```

# additional parameters 

You can play a bit with changing your function to something that you can't integrate "by hand"

BUT we might want more parameters

to work with ODE,
parameters must all be input as a single list; similar to how we return multiple outputs from a function

see example below..lets add a carrying capacity

# R code with carrying capacity

```{r}
# create parameter list
initialrabbits <- 20
years <- seq(from = 1, to = 100, by = 2)
parms <- c(0.01)

# apply solver
results <- ode(y = initialrabbits, times = years, 
              func = dexppop, parms = parms)
head(results)
```


```{r}
# add more meaningful names
colnames(results) <- c("year","P")

# plot
ggplot(as.data.frame(results), 
       aes(x = year, 
           y = P)) + 
  geom_point() + 
  labs(y = "Population", 
       x = "years")
```


```{r}
# try again with different parameters
alternativeparms <- c(0.05)

results_2 <- ode(y = initialrabbits, times = years, 
                func = dexppop, parms = alternativeparms)

# look at results
head(results_2)
```


```{r}
colnames(results_2) = c("year","P_parm2")

# plot
ggplot(as.data.frame(results_2), 
       aes(x = year, 
           y = P_parm2)) + 
  geom_point() + 
  labs(y = "Population", 
       x = "years")
```


```{r}
# compare by combining into a single data frame
both <- inner_join(x = as.data.frame(results), 
                   y = as.data.frame(results_2))

both_p <- both %>% 
  pivot_longer(cols = 2:3,
               names_to = "model", 
               values_to = "P")
  #gather(key = "model", value = "P", -year) # could use pivot_longer instead of gather

ggplot(both_p, 
       aes(x = year,
           y = P, 
           col = model)) + 
  geom_point() + 
  labs(y = "Population", "years")

# try playing on your own - modify the function in some way
```

# Difference Equations 

What is ODE doing? (iterating in 'smart ways')

Similar to "difference equations"

## Difference Equations

Population models can be discrete (rather than continuous)
So we could implement them as difference equations and iterate


```{r iteratie}
# how many rabbits after 50 years given a growth of 0.1
# starting with 1 rabbit - but a carrying capacity of 500

# lets also keep the parameters for use later
P0 <- 1
r <- 0.05
K <- 200
T <- 50

discrete_result <- discrete_logistic_pop(P0 = 1, r = 0.05, K = 200, T = 50)
discrete_result
```


# Differential Equation, Difference (Iteration by hand) comparison

Remember we have 3 ways now to calculate population

analytical solution - based on integration (exppop.R) BEST

using an ode solver for numerical approximation (exppop_play.R)

numerical integration using in discrete steps (discrete_logistic_pop.R)
```{r closed}
analytic_result <- exppop(P0 = P0, r = r, K = K, T = T)

analytic_result
discrete_result
# why are they different
# look at trajectories
```


```{r closed}
growth_result <- data.frame(time = seq(from = 1, to = 100))

growth_result$Panalytic <- 
  growth_result$time %>% 
  map_dbl(~exppop( P0 = 1, r = 0.05, K = 200, T = .x ))

growth_result$Pdiscrete <- 
  growth_result$time %>% 
  map_dbl(~discrete_logistic_pop( P0 = 1, r = 0.05, K = 200, T = .x ))

tmp <- 
  growth_result %>% 
  gather(key = "Ptype", value = "P", -time)

ggplot(data = tmp, 
       aes(x = time, 
           y = P, 
           col = Ptype)) + 
  geom_point()

# try running them for longer time periods to see what happens 
# change the value of r, K , P0 - see how it effects the results
```

# Compare analytical, difference and ODE

Finally look at continuous derivative using ODE solve
Needs
*initial condtions
*differential equation
*parameters

```{r}
# set up using the same parameters
pcompare <- list(r = r, carry_capacity = K)

# now run our ODE solver
result <- ode(P0, growth_result$time, dexppop_play, pcompare)
head(result)
```


```{r}
# we already have time - so just extract population
growth_result$Pdifferential <- result[,2]

# comapre all 3 approaches
tmp <- 
  growth_result %>% 
  gather(key = "Ptype",value="P",-time)

ggplot(data = tmp, 
       aes(x = time, 
           y = P, 
           col = Ptype)) + 
  geom_point()

# notice Pdifferential is closer to Panalytic than Pdiscrete
```





