---
title: "sobol_with_ode"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)

source("dexppop_play.R")
```


```{r sen}
# want to learn about sensitivity to growth rate (r) and carrying capacity 

# we will assume that we know the initial population,
Pinitial <- 10

# set the number of parameters
np <- 100

# get our 1st set of samples
carry_capacity <- rnorm(mean = 400, sd = 50, n = np)
r <- rnorm(mean = 0.01, sd = 0.005, n = np)
X1 <- cbind.data.frame(r = r, carry_capacity = carry_capacity)

# repeat to get our second set of samples
carry_capacity = rnorm(mean = 400, sd = 50, n = np)
r <- rnorm(mean = 0.02, sd = 0.005, n = np)
X2 <- cbind.data.frame(r = r, carry_capacity = carry_capacity)

# create our sobel object and get sets of parameters for running the model
sens_P <- sobolSalt(model = NULL,X1, X2, nboot = 300)

# our parameter sets are
head(sens_P$X)
```


```{r sen}
# lets add names to columns
colnames(sens_P$X) <- c("r","carry_capacity")

# run our differential equation and keep the output
# BUT
# what output do we want  to keep
# how about maximum population if we run the model for 200 years, and how many years to get to the carrying capacity

# for illustration lets look at running just one parameter sets and summarizing results
sens_P$X[1,]
# recall ODE needs ALL of our parameters in a single list 
# initial population and times for which we want output 

# gets results for 200 years (evaluating every year)
simtimes <- seq(from = 1, to = 200)
parms <- list(r = sens_P$X[1, "r"], 
             carry_capacity = sens_P$X[1, "carry_capacity"])

# or
parms <- list(r = as.data.frame(sens_P$X)$r[1], 
             carry_capacity = as.data.frame(sens_P$X)$carry_capacity[1])

result <- ode(y = Pinitial, times = simtimes, 
             func = dexppop_play, parms = parms)

head(result)
```


```{r sen}
colnames(result) <- c("time","P")
# turn it into a data frame
result <- as.data.frame(result)
ggplot(result, aes(time, P)) + 
  geom_point()
```


```{r sen}
# extra our metrics of interest  from this
# maximum population it gets to
maxpop <- max(result$P)
maxpop
```


```{r sen}
# years required to get to the maximum poputation
# which.max will tell us when max population occur
idx <- which.max(result$P)
# turn this index into a year (might be the same if time step in 1 but just in case it isn't)
maxyear <- result$time[idx]
maxyear
```

# Compute our metric for all the parameter sets

What if we want to run for all parameters

Lets create two additional functions that will help us

* a function that computes the metrics we want

* a function that runs our ode solver and computes the metrics (I call it a wrapper function as it is really just a workflow/wrapper to call ode solver and then compute metrics)
```{r sen2}
## -- turn computing our metrics into a function -- ##

compute_metrics <- function(result) {
  maxpop <- max(result$P)
  idx <- which.max(result$P)
  maxyear <- result$time[idx]
  return(list(maxpop = maxpop, maxyear = maxyear))
}

# try it on our first parameter set
compute_metrics(result)

# great but we need to apply the ode and this function for all of our parameters
```


```{r sen2}
## -- define a wrapper function to do everything we need -- ##
# run solver and compute metrics then send back results for each parameter

p_wrapper <- function(r,carry_capacity, Pinitial, simtimes, func) {
  parms <- list(r = r, carry_capacity = carry_capacity)
  result <- ode(y = Pinitial, times = simtimes, 
                func = func, parms = parms, method = "daspk") 
  colnames(result) <- c("time","P")
  # get metrics
  metrics <- compute_metrics(as.data.frame(result))
  
  return(metrics)
}

# now use pmap to run wrapper function for each parameter set
allresults <- as.data.frame(sens_P$X) %>% 
  pmap(p_wrapper, Pinitial = Pinitial, 
       simtimes = simtimes, func = dexppop_play)

# extract out results from pmap into a data frame
allres <- allresults %>% 
  map_dfr(`[`,c("maxpop","maxyear"))


# create boxplots
tmp <- allres %>% 
  pivot_longer(cols = 1:2, 
               names_to = "metric",
               values_to = "value")

ggplot(tmp, 
       aes(x = metric, 
           y = value, 
           col = metric)) + 
  geom_boxplot()
```

Compute the sobol indicies for each metric

```{r sen3}
# sobol can only handle one output at a time  - so we will need to do them separately
sens_P_maxpop <- sensitivity::tell(sens_P, allres$maxpop)

# first-order indices (main effect without co-variance)
sens_P_maxpop$S
```


```{r sen3}
# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_P_maxpop$T
```



```{r sen3}
# create another one for max year
sens_P_maxyear <- sensitivity::tell(sens_P, allres$maxyear)

# first-order indices (main effect without co-variance)
sens_P_maxyear$S
```


```{r sen3}
# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_P_maxyear$T
```



